---
title: 'Exp 3 Results: Plant Isotopes'
author: "Yaya"
date: "19 August 2021"
output: 
  html_document: 
    toc: yes
    toc_depth: 4
    toc_float: yes
    theme: united
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggpubr)
library(plotly)
library(viridis)
library(lmerTest)
library(emmeans)
library(multcomp)
library(lme4)
library(tidyverse)
library(janitor)
set.seed(20210918)

source(here::here('src/yaya_fxns.R'))
source("C:/Users/yaya/Dropbox/1 Ecologist/2 EXPERIMENTS/yaya_r_themes.R")

plot_levels <- c('ghop', 'mantid', 'spider', 'control')
model_levels <- c('control', 'ghop', 'mantid', 'spider')

theme_npp <- function() {
  theme_bw() + 
    theme(axis.text.x = element_text(vjust = 1, hjust =0.5),
          panel.background = element_rect(fill = "#f9f9f9",
                                          colour = "#f9f9f9"),
          panel.border = element_rect(color = "black", fill = NA)) +
    theme(plot.title = element_text(hjust = 0.5))
  
}
```


```{r echo=FALSE}
## import cage meta data
cages0 <- read_csv(here::here('data/0_cage_allocation.csv')) %>% 
  select(-predatorID) %>% 
  filter(cage != 3)


cages0$block <- as_factor(cages0$block)
cages0

# data import updated 2021 feb, more simple/streamlined
npp_raw <- read_csv(here::here('data/biomass_DATA_master_2021.02.csv'))
colnames(npp_raw) <- c('cage', 'fxl_group', 'm_group')
npp_raw[which(npp_raw$fxl_group == 'C. argentatus'),]$fxl_group <- 'C.arg'


```

## Total plant biomass

xxxxx

These are graphs printed in succession (can be two in a row, working on four in a table)

```{r echo=FALSE}
## CREATING THE MASTER DATA FRAME ####
# add all functional group masses by cage
plant_npp <- npp_raw %>% 
  na.omit() %>% 
  group_by(cage, fxl_group) %>% 
  summarize(cage = first(cage), 
            fxl_group = first(fxl_group), 
            m_group = sum(m_group)) %>% 
  ungroup() %>% 
  left_join(cages0) %>% 
  select(cage, block, position, treatment, replicate, everything()) %>%
  filter(cage != 40, # the cage with MASSIVE amounts of legumes
         cage != 26) %>%  #the cage where the mantid died very early
  mutate_if(is.numeric, round, 4) %>%
  # pivot_longer(names_to = 'fxl_group', values_to = 'm_group', C.arg:other) %>% 
  mutate(treatment = as_factor(treatment),
         treatment = fct_relevel(treatment, c('ghop', 'mantid', 'spider', 'control'))) %>%
  arrange(cage, fxl_group)

## for graphs by cage
plant_npp_cage <- plant_npp %>% 
  group_by(cage) %>% 
  summarize(cage = first(cage),
            block = first(block),
            treatment = first(treatment),
            cage_npp = sum(m_group))

## for graph by percent
plant_npp_percent <- plant_npp %>% 
  left_join(plant_npp_cage) %>% 
  mutate(percent_cage_npp = m_group / cage_npp) %>% 
  select(-cage_npp)

```

This is a visualization of the total productivity in each cage.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Pre-check: visualizing cage vs npp ####
npp_cage <- plant_npp_cage %>% 
  ggplot(aes(x = block, 
             y = cage_npp, 
             group = cage,
             fill = treatment,
             text = paste("cage:", cage))) +
  geom_col(position = position_dodge(),
           color = 'black') +
  scale_fill_viridis(name = 'Treatment', 
                     # labels = c('Grasshopper', 'Mantid', 'Spider', 'Litter-only\ncontrol'), 
                     discrete = TRUE) +
  labs(title = 'Total plant biomass by cage',
       x = 'Cage block',
       y = 'Total biomass (g)',
       color = 'Treatment') +
  theme_npp()

# ggsave(here::here('results/plant_productivity_by.block.png'),width = 7, height = 5, dpi = 600)

```

### Overall plant productivity, by cage

How does the total plant biomass differ between cages? 

```{r}

## ..Figure: overall plant productivity -----
plant_npp_cage %>%
  group_by(treatment) %>% 
  mutate(treatment = fct_relevel(treatment, plot_levels)) %>% 
  summarize(mean_cage_npp = mean(cage_npp),
            se_cage_npp = se(cage_npp)) %>% 
  ggplot(aes(x = treatment, 
             y = mean_cage_npp, 
             fill = treatment)) +
  geom_col(color = 'black') +
  geom_errorbar(aes( ymin = mean_cage_npp - se_cage_npp, 
                     ymax = mean_cage_npp + se_cage_npp),
                width = 0.2) +
  geom_jitter(data = plant_npp_cage, 
              aes(x = treatment, 
                  y = cage_npp),
              width = 0.2) + 
  scale_fill_viridis(name = '', discrete = T) +
  labs(title = 'Cage plant productivity by treatment',
       x = 'Treatment',
       y = 'Total plant biomass (g)',
       color = 'Treatment') +
  scale_x_discrete(labels=c('Grasshopper', 'Mantid', 'Spider', 'Control')) +
  theme_yaya() + 
  theme(legend.position = 'none',
        plot.title = element_blank())

# my_ggsave(here::here('results/7.2_plant_productivity_trt.png'))

# ggsave(here::here('results/7.2_plant_productivity_trt.png'),
# width = 5.5, height = 5, dpi = 600)
```

And the statistical analysis:

```{r}
## Data analysis: cage npp by treatment----
## relevels treatment so model comparisons are against the control
plant_npp_cage$treatment <- fct_relevel(plant_npp_cage$treatment, model_levels)

lmm_npp_cage <- lmer(cage_npp ~ treatment + (1|block), data = plant_npp_cage)
summary(lmm_npp_cage) ## non significant

plot(lmm_npp_cage)

# This is for model selection, not for looking at factors within the model
anova(lmm_npp_cage)
rand(lmm_npp_cage)

## Not running this section because no effects were significant
# # running emmeans ("estimating marginal means") as an example, no significance found above
# lmm_npp_cage.emm <- emmeans(lmm_npp_cage, "treatment")
# pairs(lmm_npp_cage.emm)
```

For the overall cage NPP, there is no difference between treatments.

#### Model assumption testing (residual distribution, homskedasticity)
To check the model validity, this follows the coding club tutorial for basic linear models:

```{r}
##=======from the Coding Club tutorial for basic linear models
# #1
# Checking that the residuals are normally distributed
# Extracting the residuals
lmm_npp_cage.resid <- resid(lmm_npp_cage)
# Using the Shapiro-Wilk test
shapiro.test(lmm_npp_cage.resid) ## significant? _________***

# The null hypothesis of normal distribution is accepted: there is no significant difference (p > 0.05) from a normal distribution

# #2
# Checking for homoscedasticity
bartlett.test(cage_npp ~ treatment, data = plant_npp_cage)  # Note that these two ways of writing the code give the same results

# The null hypothesis of homoscedasticity is ___ accepted/rejected***

```

### Plant productivity by functional group

Here's the spread of plant productivity by functional group, by treatment + cage, and by percentage of each cage's productivity.

```{r}
par(mar = c(4, 4, .1, .1))

## ..Figure:  plant mass by fxl group -----
# simple histogram for the distribution of masses by fxl group
hist(plant_npp$m_group)

## .. Figure: plant mass by functional group and cage
plant_npp %>% 
  ggplot(aes(x = cage, 
             y = m_group, 
             group = block,
             fill = fxl_group,
             text = paste("cage:", cage))) +
  geom_col(position = position_dodge(),
           color = 'black') +
  scale_fill_viridis(name = 'Treatment', 
                     # labels = c('Grasshopper', 'Mantid', 'Spider', 'Litter-only\ncontrol'), 
                     discrete = TRUE) +
  labs(title = 'Total plant biomass by cage',
       x = 'Cage',
       y = 'Total biomass (g)',
       color = 'Treatment') +
  facet_wrap(vars(treatment))+
  theme_npp()


### ..Figure: stacked barchart of fxl group  %NPP by treatment ####
plant_npp_percent %>% 
  # mutate(cage = as_factor(cage)) %>% 
  ggplot(aes(cage, 
             # percent_cage_npp,
             m_group,
             # group = treatment,
             fill = fxl_group))+
  geom_col(position = "fill") +
  scale_y_continuous(labels = scales::percent) + 
  scale_fill_viridis(discrete = T, name = 'Functional group') +
  labs(x = "Cage",
       y = "Percent NPP by fxnl group",
       title = "Plant functional group %NPP, per cage") + 
  theme_npp() +
  theme(axis.text.x=element_blank(),
        axis.ticks = element_blank()) + 
  facet_grid(. ~ treatment)
```

Most group masses were under 10g. 





```{r}
plant_npp %>% 
  mutate(treatment = fct_relevel(treatment, plot_levels)) %>%
  group_by(treatment, fxl_group) %>% 
  summarize(mean_m_group = mean(m_group),
            se_m_group = se(m_group)) %>% 
  # View()
  ggplot(aes(treatment, 
             mean_m_group, 
             fill = fxl_group))+
  geom_col(position = position_dodge(),
           color = 'black') +
  geom_errorbar(aes( ymin = mean_m_group - se_m_group,
                     ymax = mean_m_group + se_m_group),
                width = 0.2,
                position = position_dodge(width = 0.9)) +
  # geom_jitter(data = plant_npp,
  #             aes(treatment, m_group, color = fxl_group),
  #             width = 0.2,
  #             position = position_jitterdodge()) +
  # geom_point(data = plant_npp, 
  #            aes(treatment, m_group, fill = fxl_group)) +
  scale_fill_viridis(name = 'Functional group', option = 'magma', discrete = TRUE) +
  labs(title = 'Plant productivity, by functional group',
       x = 'Treatment',
       y = 'Total plant biomass (g)',
       color = 'Treatment') +
  scale_x_discrete(labels=c('Grasshopper', 'Mantid', 'Spider', 'Control')) +
  theme_yaya() + 
  theme(legend.position = 'none',
        plot.title = element_blank())


# my_ggsave(here::here('results/7.2_plant_productivity_trt_fxl_group.png'))

```

### Data analyses: linear mixed models

This is the main model for consideration. 

```{r}
## Data analysis: functional group productivity by treatment----
plant_npp$treatment <- fct_relevel(plant_npp$treatment, model_levels)

## using / for nested random effects, or :
## removed the cage random factor because it doesn't make sense and there's singularity issues

lmm_fxl_npp0 <- lmer(m_group ~ treatment + fxl_group + (1|block), 
                     data = plant_npp)
summary(lmm_fxl_npp0) ## basically this result confirms the fact that each fxl group generate different biomasses. not relevant.

lmm_fxl_npp <- lmer(m_group ~ treatment * fxl_group + (1|block), data = plant_npp)
summary(lmm_fxl_npp)

#**** 
anova(lmm_fxl_npp0, lmm_fxl_npp)
#****
# what does this mean? is it even relevant?^^^^

anova(lmm_fxl_npp) # not necessary bc there is interaction significance
rand(lmm_fxl_npp)

### comparing between treatments in model without interactions
lmm_fxl_npp0.emm <- emmeans(lmm_fxl_npp0, "fxl_group")
pairs(lmm_fxl_npp0.emm)



## results say misleading due to interactions
lmm_fxl_npp.emm <- emmeans(lmm_fxl_npp, "fxl_group")
pairs(lmm_fxl_npp.emm)


### example for using emmeans
pigs.lm <- lm(log(conc) ~ source + factor(percent), data = pigs)
pigs.emm.s <- emmeans(pigs.lm, "source")
pairs(pigs.emm.s)


```

### By each functional group

Now these are the figures and regression models for each individual plant functional group. 

```{r}
## ..Figure: C arg productivity by treatment----

(npp_c.arg <- plant_npp %>% 
   filter(fxl_group == 'C.arg') %>% 
   group_by(treatment) %>% 
   mutate(treatment = fct_relevel(treatment, c('ghop','mantid', 'spider', 'control'))) %>%      # View()
   summarize(mean_m_group = mean(m_group),
             se_m_group = se(m_group),
             treatment = first(treatment)) %>% 
   ggplot(aes(treatment, 
              mean_m_group,
              fill = treatment))+
   geom_bar(position = 'dodge', stat = 'identity') + 
   geom_errorbar(aes(ymin = mean_m_group - se_m_group, 
                     ymax = mean_m_group + se_m_group), 
                 width=.2,
                 position=position_dodge(.9)) +
   scale_fill_viridis(discrete = T, name = 'Functional group') +
   labs(title = "Plant productivity by functional group: C. argentatus",
        x = 'Treatment',
        y = 'Total plant biomass (g)',
        color = 'Treatment') +
   scale_x_discrete(labels=c('Grasshopper', 'Mantid', 'Spider', 'Control')) +
   theme_yaya() + 
   theme(legend.position = 'none',
         plot.title = element_blank())
)

# my_ggsave(here::here('results/7.2_plant_productivity_C.arg.png'))
```


```{r}
## Data analysis: linear model for C.arg npp ####
plant_npp$treatment <- fct_relevel(plant_npp$treatment, model_levels)

lmm_npp_c.arg <- lmer(m_group ~ treatment + (1|block),
                      filter(plant_npp, fxl_group == 'C.arg'))
summary(lmm_npp_c.arg)
emmeans(lmm_npp_c.arg, pairwise ~ treatment) #method 1


anova(lmm_npp_c.arg)
rand(lmm_npp_c.arg)
plot(lmm_npp_c.arg)
```



```{r}
## ..Figure: cereals productivity by treatment----

(npp_cereals <- plant_npp %>% 
   group_by(treatment) %>% 
   filter(fxl_group == 'cereals') %>% 
   mutate(treatment = fct_relevel(treatment, c('ghop','mantid', 'spider', 'control'))) %>% 
   # View()
   summarize(mean_m_group = mean(m_group),
             se_m_group = se(m_group),
             treatment = first(treatment)) %>% 
   ggplot(aes(treatment, 
              mean_m_group,
              fill = treatment))+
   geom_bar(position = 'dodge', stat = 'identity') + 
   geom_errorbar(aes(ymin = mean_m_group - se_m_group, 
                     ymax = mean_m_group + se_m_group), 
                 width=.2,
                 position=position_dodge(.9)) +
   scale_fill_viridis(discrete = T, name = 'Functional group') +
   
   labs(x = "Treatment",
        y = "Total biomass, g",
        title = "Plant productivity by functional group: cereals") + 
   scale_x_discrete(labels=c('Grasshopper', 'Mantid', 'Spider', 'Litter-only\ncontrol')) +
   theme_npp()
)
# npp_cereals
```


```{r}
## Data analysis: linear model for cereals npp ####
plant_npp$treatment <- fct_relevel(plant_npp$treatment, model_levels)

lmm_npp_cereals <- lme4::lmer(m_group ~ treatment + (1|block),
                              filter(plant_npp, fxl_group == 'cereals'))
summary(lmm_npp_cereals) ## ghop and mantid different from controls


lmm_npp_cereals2 <- lmerTest::lmer(m_group ~ treatment + (1|block),
                                   filter(plant_npp, fxl_group == 'cereals'))
summary(lmm_npp_cereals2) ## ghop and mantid different from 

## so check the difference between levels
# method 1: emmeans "estimating marginal means" aka least squares means
emmeans(lmm_npp_cereals, pairwise ~ treatment, lmer.df="satterthwaite") #method 1
emmeans(lmm_npp_cereals2, pairwise ~ treatment) #method 1


# running glht(), method 2
post.hoc <- glht(lmm_npp_cereals, linfct = mcp(treatment = 'Tukey'))
# displaying the result table with summary()
summary(post.hoc)

anova(lmm_npp_cereals)
rand(lmm_npp_cereals)
plot(lmm_npp_cereals)


## nevo method!
lmm_npp_cereals_null <- lme4::lmer(m_group ~ (1|block),
                                   filter(plant_npp, fxl_group == 'cereals'))
anova(lmm_npp_cereals)
anova(lmm_npp_cereals_null, lmm_npp_cereals)
emmeans(lmm_npp_cereals, pairwise ~ treatment)
```


```{r}
## ..Figure: legumes productivity by treatment----
(npp_legumes <- plant_npp %>% 
   group_by(treatment) %>% 
   filter(fxl_group == 'legumes') %>% 
   mutate(treatment = fct_relevel(treatment, c('ghop','mantid', 'spider', 'control'))) %>% 
   # View()
   summarize(mean_m_group = mean(m_group),
             se_m_group = se(m_group),
             treatment = first(treatment)) %>% 
   ggplot(aes(treatment, 
              mean_m_group,
              fill = treatment))+
   geom_bar(position = 'dodge', stat = 'identity') + 
   geom_errorbar(aes(ymin = mean_m_group - se_m_group, 
                     ymax = mean_m_group + se_m_group), 
                 width=.2,
                 position=position_dodge(.9)) +
   scale_fill_viridis(discrete = T, name = 'Functional group') +
   
   labs(x = "Treatment",
        y = "Total biomass, g",
        title = "Plant productivity by functional group: legumes") + 
   # scale_x_discrete(labels=c('Grasshopper', 'Mantid', 'Spider', 'Litter-only\ncontrol')) +
   theme_npp()
)
```


```{r}
## Data analysis: linear model for legumes npp -----
plant_npp$treatment <- fct_relevel(plant_npp$treatment, model_levels)

lmm_npp_legumes <- lmer(m_group ~ treatment + (1|block),
                        filter(plant_npp, fxl_group == 'legumes'))
summary(lmm_npp_legumes)
plot(lmm_npp_legumes)

emmeans(lmm_npp_legumes, pairwise ~ treatment, lmer.df="satterthwaite") #method 1
```


```{r}
## ..Figure: other productivity by treatment----

(npp_other <- plant_npp %>% 
   group_by(treatment) %>% 
   filter(fxl_group == 'other') %>% 
   mutate(treatment = fct_relevel(treatment, c('ghop','mantid', 'spider', 'control'))) %>% 
   # View()
   summarize(mean_m_group = mean(m_group),
             se_m_group = se(m_group),
             treatment = first(treatment)) %>% 
   ggplot(aes(treatment, 
              mean_m_group,
              fill = treatment))+
   geom_bar(position = 'dodge', stat = 'identity') + 
   geom_errorbar(aes(ymin = mean_m_group - se_m_group, 
                     ymax = mean_m_group + se_m_group), 
                 width=.2,
                 position=position_dodge(.9)) +
   scale_fill_viridis(discrete = T, name = 'Functional group') +
   
   labs(x = "Treatment",
        y = "Total biomass, g",
        title = "Plant productivity by functional group: other") + 
   scale_x_discrete(labels=c('Grasshopper', 'Mantid', 'Spider', 'Litter-only\ncontrol')) +
   theme_npp()
)
```


```{r}
## Data analysis: linear model for other npp ####
plant_npp$treatment <- fct_relevel(plant_npp$treatment, model_levels)

lmm_npp_other <- lmer(m_group ~ treatment + (1|block),
                      filter(plant_npp, fxl_group == 'other'))
summary(lmm_npp_other)
emmeans(lmm_npp_other, pairwise ~ treatment) #method 1
post.hoc <- glht(lmm_npp_other, linfct = mcp(treatment = 'Tukey')) # method 2
summary(post.hoc)
```


```{r}
####..Figure: plant dominance (aka percent of total biomass) .. ####

### nothing yet


```


```{r}


```

