---
title: "Chapter 3: feeding assay results"
author: "Yaya"
date: "15 August 2021"
output: 
  html_document: 
    toc: yes
    toc_depth: 4
    toc_float: yes
    theme: united
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	include = FALSE
)
library(tidyverse)
library(here)
library(readr)
library(janitor)
library(viridis)
library(gridExtra)
source(here::here('src/yaya_fxns.R'))

vir_4 <- viridis(4)

# predators to remove bc they didn't eat
remove_predators <- tibble(predator_id = c('MR08', 'S07', 'S09', 'S23'))

assay_raw <- read_csv('data/assay_feeding4_predators.csv', lazy = FALSE) %>% 
  clean_names() %>% 
  remove_empty() %>% 
  anti_join(remove_predators) %>% 
  rename(tube_feces_mg = feces_tube_feces)

calib_raw <- read_csv(here::here('data/assay_feeding4_ghops.csv')) %>% 
  clean_names() %>% 
  remove_empty() %>% 
  filter(!is.na(analytical_fate)) %>% 
  rename(ghop_calc_mass = calculated_ghop_mass,
         tube_poop_mass = tube_mas_with_poop)
```

Feeding assay #4 is to parameterize the quality and quantity of inputs to mesocosms. This is for parameterizing the predator products produced while running the experiment. 

## Calibration grasshoppers to infer predators' food

```{r}
calib_data <-  calib_raw %>% 
  mutate(ghop_wet_mass_mg  = ifelse(is.na(tube_poop_mass), 
                                    (ghop_calc_mass - tube_mass)*1000,
                                    (ghop_calc_mass - tube_poop_mass)*1000)) %>% 
  select(ghop_id, ghop_wet_mass_mg, analytical_fate, predator_id, comments)

```

**STILL NEEDS** dry ghop mass data (check datasheets)


## Predator products, quantity

This is an R markdown document for calculating and summarizing the predator products assay for quantifying the amount of necromass entering the mesocosms in chapter 3.


```{r calculate_product_masses}
assay_wide_data <- assay_raw %>% 
  select(-feeding_observations, -original_id, -eggsac_notes) %>% 
  mutate(lid_mass_mg = lid_mass_g*1000,
         lid_predator_start_mg = lid_predator_start_g*1000,
         lid_predator_postprandial_mg = lid_predator_24_48_hr_g*1000,
         remains_mg = remains_tube_dry_remains_mg - remains_tube_mg,
         feces_mg = feces_tube_dry_feces_mg - feces_tube_mass_mg,
         predator_start_mass_mg = lid_predator_start_mg - lid_mass_mg,
         predator_postprandial_mass_mg = lid_predator_postprandial_mg - lid_mass_mg) %>% 
  select(predator_type, predator_id, remains_mg, feces_mg, predator_start_mass_mg, predator_postprandial_mass_mg)

```


#### Plot of predator by-products

Plot the predator mass versus their product mass to see whether there any confounding variables there.


```{r}
pairs(select(assay_wide_data, -c(predator_type, predator_id)))
```

Figures to check whether predator mass had an effect on the amount of products they produced.
```{r, figures-side, fig.show = "hold", out.width="50%"}
# pred_feces <- 
assay_wide_data %>% 
  ggplot(aes(predator_start_mass_mg, 
             feces_mg, 
             color = predator_type)) + 
  geom_point() + 
  scale_color_manual(values = viridis(2), name = 'Predator') +
  expand_limits(x = 0, y=0) + 
  labs(x = 'Predator mass at start (mg)',
       y = 'Product mass (mg)',
       title = 'Predator start mass vs feces') +
  theme_flat

# pred_remains <- 
assay_wide_data %>% 
  ggplot(aes(predator_start_mass_mg, 
             remains_mg, 
             color = predator_type)) + 
  geom_point() + 
  scale_color_manual(values = viridis(2), name = 'Predator') +
  expand_limits(x = 0, y=0) + 
  labs(x = 'Predator mass at start (mg)',
       y = 'Product mass (mg)',
       title = 'Predator start mass vs remains') +
  # theme(legend.title = element_blank()) + 
  theme_flat

# grid.arrange(pred_feces, pred_remains, ncol=2)
```


```{r}
## Not showing this--- ANOVA for predator type and start mass as factors to predict products mass
pred_mass_mod <- aov(feces_mg ~ predator_type + predator_start_mass_mg,
                 assay_wide_data)
summary(pred_mass_mod)

```


### Plot predators by feeding strategy vs foraging by-products

```{r}
vars_to_keep <- c('feces', 'remains', 'predator_start_mass', 'predator_postprandial_mass')

assay_long_data <- assay_wide_data %>% 
  pivot_longer(-c(predator_type, predator_id),
               names_pattern = '(.*)_mg',
               names_to = 'product', 
               values_to = 'mass') %>% 
  filter(product %in% vars_to_keep) %>% 
  mutate(product = factor(product, levels = c('feces', 'remains', 'predator_start_mass', 'predator_postprandial_mass')),
         predator_type = factor(predator_type, levels = c('mantid', 'spider'))) %>% 
  group_by(predator_type, product)

assay_long_data %>% 
  filter(product %in% c('feces', 'remains')) %>% 
  group_by(predator_type, product) %>% 
  summarize(mean_mass = mean(mass), 
            se_mass = se(mass)) %>% 
  ggplot(aes(product, 
             mean_mass, 
             fill = predator_type)) + 
  geom_col(position = 'dodge') + 
  geom_errorbar(aes(ymin = mean_mass - se_mass,
                    ymax = mean_mass + se_mass),
                position = position_dodge(width = 0.9),
                width = 0.15) +
  geom_point(data = filter(assay_long_data, product %in% c('feces', 'remains')),
             aes(product, 
                 mass),
             position = position_jitterdodge(dodge.width=0.9)) + 
  scale_fill_manual(values = viridis(2), name = 'Predator') +
  labs(x = 'Necromass product',
       y = 'Dry mass (mg)',
       title = 'Mantid + spider feces + waste quantities') +
  theme_flat


# summarize data

assay_long_data %>% 
  filter(product %in% c('feces', 'remains')) %>% 
  group_by(predator_type, product) %>% 
  summarize(mean_mass = mean(mass), 
            se_mass = se(mass),
            n = n()) %>% 
  knitr::kable()
```



## Elemental analysis data

Import elemental analysis

```{r}
# this is data from the elemental analysis of the assay
# both the calibration ghops and the predator products

## read data from two plates
e_plate1 <- read_excel(here::here('data/siel data/YT-F4.1_CN_Rep, SIEL SO#1768.xls'), skip = 2) %>% 
  clean_names() %>% 
  rename(mass_mg = weight_mg)

e_plate2 <- read_excel(here::here('data/siel data/YT-F4.2-PPg_CN_Rep, SIEL SO#1768.xls'), skip = 2) %>% 
  clean_names() %>% 
  rename(mass_mg = weight_mg)

# join two plates and remove instrumentation samples
e_0all <- full_join(e_plate1, e_plate2) %>% 
  filter(sample_id != 'spinach',
         sample_id != 'MR08-R') %>% # remove MR08-R because it didn't eat during assay
  mutate(exp_id = substr(sample_id, 1,1))
```

Splitting the SIEL plates' data into subsets of data frames according to the product type

(Can't really infer individualized product quantity with only one data point)

```{r}

# split dataset into subsets according to the groups above
e_preds <- e_0all %>% 
  filter(exp_id == 'M' | exp_id == 'S') %>% 
  mutate(prod_type = 'predator prods',
         exp_stage = 'assay products',
         predator_id = sub("\\-.*", "", sample_id),
         sample_type = substrRight(sample_id, 1))
e_preds$sample_type[e_preds$sample_type == 'F'] <- 'feces'
e_preds$sample_type[e_preds$sample_type == 'R'] <- 'prey remains'
e_preds$sample_type[e_preds$sample_type == 'S'] <- 'silk'
e_preds$sample_type <- as.factor(e_preds$sample_type)

e_ghops_assay <- e_0all %>% 
  filter(exp_id == 'F') %>% 
  mutate(prod_type = 'calibration ghops',
         exp_stage = 'assay prep',
         ghop_id = substrRight(sample_id, 3),
         sample_type = 'ghop')

e_ghops_mesocosm <- e_0all %>% 
  filter(exp_id == 'G') %>% 
  mutate(prod_type = 'calib ghops',
         exp_name = 'exp samples',
         ghop_num = substr(sample_id, 1, 3),
         week_num = substr(sample_id, 5, nchar(sample_id)),
         sample_type = 'ghop')

```

Now we've got three datasets for the three types of physical samples we have in our hands: assay calibration prey, assay predator products, experimental calibration prey

## UNTESTED CODE FOR MERGING GHOP DATA

```{r}

#=====
# calculate elemental composition of feeding 4 calibration ghops

# 1. merge elemental data with earlier data
ghops_all <- left_join(e_ghops, a_ghops1)

# 2. summarize data
summary(ghops_all[c('mass_mg', 'C_N_ratio')])

#====
# calculate elemental composition of feed4 assay predator products
# 1. melt lab data to have prey remains + feces data separately
# a_preds0 <- pivot_longer()

# 2. merge elemental data with lab data
preds_all <- left_join(e_preds, a_preds3)

fact_cols <- c('sample_type', 'prod_type', 'predator_type', 'expID')
preds_plots <- preds_all %>% 
  mutate_at(fact_cols, list(~ factor(.))) %>% 
  filter(sample_type != 'silk') %>% 
  group_by(predator_type, sample_type) %>% 
  ggplot(aes(x = sample_type,
             y = C_N_ratio,
             group = predatorID)) + 
  geom_boxplot() +
  facet_grid(. ~ predator_type) + 
  theme_flat

ggplotly(preds_plots)

# 
# preds_mantid <- filter(preds_all, predator_type == 'mantid')
# preds_spider <- filter(preds_all, predator_type == 'spider')
# 
# # this is elemental data only, not for mass
# ggboxplot(preds_mantid, x = 'sample_type', y = 'C_N_ratio') 
# ggboxplot(preds_spider, x = 'sample_type', y = 'C_N_ratio')
# 
# 3. summarize + viz
# preds_summ <- summarize(preds_all[c('C_N_ratio')])
# ????????????????????????????????????????????????????????????
```