---
title: 'Exp 3 Results: Soil Sample Analysis'
author: "Yaya"
date: "21 September 2021"
output: 
  html_document: 
    toc: true
    toc_depth: 4
    toc_float: true
    theme: united
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warnings = FALSE)

library(tidyverse)
library(janitor)
library(viridis)
library(plotly)
library(lme4)
library(emmeans)
library(gridExtra) # grid.arrange
library(GGally) #for ggpairs
source(here::here('src/yaya_fxns.R'))

set.seed(20210921)

## Set up my own themes
theme_angled <- theme(axis.text.x = element_text(angle = 30, vjust = 1, hjust=1),
                      panel.background = element_rect(fill = "#f9f9f9",
                                                      colour = "#f9f9f9"),
                      panel.border = element_rect(color = "black", fill = NA))

theme_flat <- theme(axis.text.x = element_text(vjust = 1, hjust =0.5),
                    panel.background = element_rect(fill = "#f9f9f9",
                                                    colour = "#f9f9f9"),
                    panel.border = element_rect(color = "black", fill = NA),
                    legend.position = "none")
```

# Soil isotope data



```{r import}
## Importing all soil isotopic data
file_names <- c(here::here('data/isotopes_litterbag_soil_plate1a_15n.csv'),
                here::here('data/isotopes_litterbag_soil_plate1a_13c.csv'),
                here::here('data/isotopes_litterbag_soil_plate1b_15n.csv'),
                here::here('data/isotopes_litterbag_soil_plate1b_13c.csv'),
                here::here('data/isotopes_litterbag_soil_plate2_15n.csv'),
                here::here('data/isotopes_litterbag_soil_plate2_13c.csv'))

# import data + clean up columns
plates <- file_names %>% 
  map_df(~read_csv(., skip=2)) %>% 
  clean_names() %>% 
  mutate(cage = substr(sample_id, 1,2)) %>% 
  rename(d13c = d13cvpdb,
         d15n = d15nair) %>% 
  select(-c(sample_wt_mg, x7, x8, x9, ampl_28_m_v, ampl_44_m_v))

# clean up rows
plates <- plates[rowSums(is.na(plates)) != ncol(plates), ] %>% 
  filter(sample_id != 'spinach',
         sample_id != '1577C-B') %>% 
  select(sample_id, cage, everything())

# to fix the cage names and turn into numeric
plates$cage <-  plates$cage %>% 
  str_replace('E', '') %>% 
  str_replace('W', '')
plates[,2:8] <- sapply(plates[,2:8],as.numeric)

plates_a <- plates %>% 
  select(cage, sample_id, total_percent_n, d15n, atom_percent_15n) %>% 
  filter(complete.cases(.))

plates_b <- plates %>% 
  select(cage, sample_id, total_percent_c, d13c, atom_percent_13c) %>% 
  filter(complete.cases(.))

plates_all <- full_join(plates_a, plates_b)

###########################
# merge with cage treatment data

## vars for releveling factors for graphs versus models
model_levels <- c('control', 'ghop', 'mantid', 'spider')
graph_levels <- c('ghop', 'mantid', 'spider', 'control')


cage_treatments <- read_csv(here::here('results/g1_updated_cages_wk9.csv')) %>% 
  clean_names() %>% 
  select(-c('predator_id', 'position', 'replicate')) %>%
  mutate(treatment = factor(treatment, levels = model_levels)) ## data default is model levels

# merge isotope data with cage meta-data
isotopes_raw <- plates_all %>% 
  left_join(cage_treatments) %>% 
  select(sample_id, cage, treatment, everything())

## two samples were doubly labeled
## these lines correctly identify them
isotopes_raw[which(isotopes_raw$sample_id == '39WU2-1'),]$sample_id <- '39WU2'
isotopes_raw[which(isotopes_raw$sample_id == '39WU2-2'),]$sample_id <- '39WB2'

# naming it after the two directions/ two samplings it is named after
isotopes_e_w_prep <- isotopes_raw %>% 
  mutate(dir = as_factor(if_else(cage>=10, substr(sample_id, 3, 3), substr(sample_id, 2, 2))),
         sample_type = if_else(cage>=10, substr(sample_id, 4, 4), substr(sample_id, 3, 3)),
         rep = if_else(cage>=10, substr(sample_id, 5, 5),substr(sample_id, 4, 4)),
         treatment_vec = as.numeric(treatment)) %>% 
  select(treatment, sample_id, sample_type, rep, everything()) %>% 
  arrange(sample_id) %>% 
  filter(sample_type == 'U')

# only for cores from under litterbags, averaging the two
isotopes_e_w <- isotopes_e_w_prep %>%
  filter(sample_type == 'U') %>%
  group_by(cage, dir) %>%
  summarise_at(vars(-c('treatment', 'sample_id', 'sample_type', 'rep')), mean) %>%
  left_join(cage_treatments) %>% 
  mutate(sampling = ifelse(dir == 'E', 2, 3))

isotopes_e_w_model <- isotopes_e_w %>% 
  mutate(treatment = factor(treatment, levels = model_levels))

isotopes_e_w_graph <- isotopes_e_w %>% 
  mutate(treatment = factor(treatment, levels = graph_levels))

```

All data imported, cleaned, and joined. Data frames available are:
isotopes_e_w (precursors: isotopes_e_w_raw, isotopes_e_w_prep)
isotopes_e_w_model
isotopes_e_w_graph


## Pivoting the dataset long and summarizing:
isotopes_under_long
isotopes_under_long_summ


```{r echo=FALSE, message=FALSE, warning=FALSE}
# lengthen dataset and summarize
isotopes_under_long <- isotopes_e_w %>% 
  ungroup() %>% 
  pivot_longer(c('total_percent_n', 'd15n', 'atom_percent_15n', 
                 'total_percent_c', 'd13c', 'atom_percent_13c'), 
               names_to = "analysis", values_to = "a_value") %>% 
  mutate(treatment = factor(treatment, levels = graph_levels))

isotopes_under_long_summ <- isotopes_under_long %>% 
  group_by(analysis, treatment) %>% 
  summarize(iso_mean = mean(a_value),
            iso_se = se(a_value),
            iso_var = var(a_value),
            iso_n = n())

# knitr::kable(isotopes_under_long_summ)
```


## Exploratory dotplots

```{r echo=FALSE, message=FALSE, warning=FALSE}
par(mfrow = c(1, 2))
dotchart(isotopes_e_w$d15n, 
         groups = isotopes_e_w$treatment_vec,
         ylab = 'treatment',
         xlab = 'd15n', main = 'cleveland dotplot: d15n',
         pch = isotopes_e_w$treatment_vec)
dotchart(isotopes_e_w$d13c, 
         groups = isotopes_e_w$treatment_vec,
         ylab = 'treatment',
         xlab = 'd13c', main = 'cleveland dotplot: d13c',
         pch = isotopes_e_w$treatment_vec)

## Pairs for all variables
# ggpairs(select(isotopes_e_w, c('total_percent_n', 'd15n','atom_percent_15n', 'total_percent_c', 'd13c','atom_percent_13c')))

```

## Histograms of all analyses

```{r echo=FALSE, message=FALSE, warning=FALSE}
##### GRAPHS, exploratory -----
# super basic histogram
# hist(isotopes_e_w$d15n)

explore_soil_hist <- function(var_name) {
  
  isotopes_under_long %>% 
    filter(analysis == var_name) %>% 
    ggplot(aes(a_value))+
    geom_histogram(bins = 10) + 
    labs(title = paste('Histogram of', var_name, 'values'),
         x = var_name) +
    # facet_grid(. ~ treatment) +
    theme_light()
}

par(mfrow = c(2,2))

explore_soil_hist('total_percent_c')
explore_soil_hist('d13c')
explore_soil_hist('total_percent_n')
explore_soil_hist('d15n')
```


### d15N histogram

Here's a histogram of the d15N data for soil samples E + W:

```{r echo=FALSE, message=FALSE, warning=FALSE}
## this is for all soil samples that were analyzed (sampling 2 + 3)
isotopes_e_w_graph %>% 
  ggplot(aes(d15n,
             fill = dir))+
  geom_histogram(bins = 10) + 
  facet_grid(dir ~ .)

## just to check, lets take a look also at the original raw data (before averaging the two samples)
isotopes_e_w_prep %>% 
  mutate(treatment = factor(treatment, levels = graph_levels)) %>% 
  ggplot(aes(block, 
             d15n,
             color = treatment)) +
  geom_point()+ 
  scale_color_viridis(discrete = TRUE) +
  labs(title = 'Original isotope data from soil samples (doubles per cage)') +
  facet_grid(dir ~ treatment)

isotopes_e_w_prep %>% 
  mutate(treatment = factor(treatment, levels = graph_levels)) %>% 
  ggplot(aes(block, 
             total_percent_n,
             color = treatment)) +
  geom_point()+ 
  scale_color_viridis(discrete = TRUE) +
  labs(title = 'Original isotope data from soil samples (doubles per cage)') +
  facet_grid(dir ~ treatment)

## same data with d15N values averaged by cage
isotopes_e_w_graph %>% 
  ggplot(aes(cage, 
             d15n,
             color = treatment)) +
  geom_point()+ 
  scale_color_viridis(discrete = TRUE) +
  labs(title = 'Averaged isotope data from soil samples (one per cage)')+
  facet_grid(dir ~ treatment)
```

#### Check z-scores for outliers

```{r echo=FALSE, message=FALSE, warning=FALSE}
## check the z scores
isotopes_e_w_graph <- isotopes_e_w_graph %>% 
  ungroup %>% 
  mutate(zscore_d15n = (d15n - mean(d15n)) / sd(d15n))
# View()

```

## Barcharts of all analyses

Take a look at all elemental analyses in one chart, in one axes.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# bar plot of all the variables, one chart

isotopes_under_long_summ %>% 
  mutate(treatment = fct_relevel(treatment, graph_levels)) %>% 
  ggplot(aes(fill = treatment, x = analysis, y = iso_mean))+
  geom_bar(position = 'dodge', stat = 'identity')+
  geom_errorbar(aes(ymin=iso_mean-iso_se, ymax= iso_mean+iso_se), width=.2,
                position=position_dodge(.9)) +
  geom_jitter(data = isotopes_under_long,
              aes(x= analysis, y = a_value), 
              position = position_dodge(.9)) +
  scale_fill_viridis(name = 'Grasshopper fate', discrete = TRUE) +
  scale_color_viridis(discrete = TRUE) +
  labs(title = 'Comparing all elemental analyses by treatment',
       x = 'Analysis',
       y = 'Mean +- se') +
  scale_x_discrete(labels=c('Atom percent 13C', 'Atom percent 15N',
                            'd13C', 'd15N',
                            'Total percent C', 'Total percent N')) +
  facet_grid(.~ dir)+
  theme_angled

# bar plot of all vars, faceted

isotopes_under_long_summ %>% 
  mutate(treatment = fct_relevel(treatment, graph_levels)) %>% 
  ggplot(aes(fill = treatment, x = treatment, y = iso_mean))+
  geom_bar(position = 'dodge', stat = 'identity')+
  geom_errorbar(aes(ymin=iso_mean-iso_se, ymax= iso_mean+iso_se), width=.2,
                position=position_dodge(.9)) +
  geom_jitter(data = isotopes_under_long,
              aes(x= treatment, y = a_value), 
              position = position_jitterdodge(.9)) +
  scale_fill_viridis(name = 'Grasshopper fate', discrete = TRUE) +
  scale_color_viridis(discrete = TRUE) +
  labs(title = 'Comparing all elemental analyses by treatment',
       x = 'Analysis',
       y = 'Mean +- se') +
  # scale_x_discrete(labels=c('Atom percent 13C', 'Atom percent 15N',
  #                           'd13C', 'd15N',
  # 'Total percent C', 'Total percent N')) +
  facet_wrap(vars(analysis), scales = 'free')
# theme_angled
```

## d15N in the soil

### boxplot
```{r echo=FALSE, message=FALSE, warning=FALSE}
###### ...Figure for d15N=====

## boxplot of d15n
(d15n_box <- isotopes_e_w_graph %>% 
   ggplot(aes(x = treatment, 
              y = d15n, 
              fill = treatment))+
   geom_boxplot() + 
   geom_jitter(width = 0.1,
               aes(color = cage)) + 
   scale_fill_viridis(name = 'Treatment', discrete = TRUE) +
   facet_grid(treatment ~ dir) +
   theme_flat)

ggplotly(d15n_box)
```

### barplots

```{r echo=FALSE, message=FALSE, warning=FALSE}
## barplot for mean d15N by treatment and direction
(soil_d15n <- isotopes_e_w_graph %>% 
   group_by(treatment) %>% 
   summarize(mean_d15n = mean(d15n),
             se_d15n = se(d15n)) %>% 
   ungroup() %>% 
   ggplot(aes(x = treatment, y = mean_d15n, fill = treatment))+
   geom_col(position = 'dodge',
            color = 'black')+
   # geom_jitter(data = filter(isotopes_e_w, dir == 'W'),
   geom_jitter(data = isotopes_e_w,
               aes(x = treatment, 
                   y = d15n,
                   group = cage),
               width = 0.2) +
   geom_errorbar(aes( ymin = mean_d15n - se_d15n, 
                      ymax = mean_d15n + se_d15n,
                      width = 0.15)) +
   scale_fill_viridis(name = 'Grashopper fate', discrete = TRUE) +
   labs(title = 'd15N of soil beneath litter bags, samplings 2 and 3',
        x = 'Treatment',
        y = 'd13c') +
   # scale_x_discrete(labels=c('Grasshopper', 'Mantid', 'Spider', 'Control')) +
   theme_flat + 
   theme(legend.position = 'none') + 
   facet_wrap(vars(dir))
)
# ggplotly(soil_d13c)

# my_ggsave(here::here('results/soil_d15n.png'))
```


### Statistical analysis

#### Sampling 2 d15N, from under East litterbags

```{r}

d15n_s2_all <- lme4::lmer(d15n ~ treatment + (1|block), filter(isotopes_e_w_model, sampling == 2))
d15n_s2_null <- lme4::lmer(d15n ~ 1 + (1|block), filter(isotopes_e_w_model, sampling == 2))


summary(d15n_s2_all)
summary(d15n_s2_null)
anova(d15n_s2_all, d15n_s2_null)

# not running emmeans because it's not significant
# d15n_s2_all.emm <- emmeans(d15n_s2_all, pairwise ~ treatment) #, lmer.df="satterthwaite") #method 1
# summary(d15n_s2_all.emm)
```
Not significant.

```{r}
(s2_d15n <- isotopes_e_w_graph %>% 
    filter(sampling == 2) %>% 
    ggplot(aes(x = cage, 
               y = d15n, 
               fill = treatment))+
    geom_boxplot() + 
    geom_jitter(width = 0.1,
                aes(color = cage)) + 
    scale_fill_viridis(name = 'Treatment', discrete = TRUE) +
    facet_grid(treatment ~ block) +
    theme_flat)
```

#### Sampling 3, from under West litterbags

```{r}
d15n_s3_all <- lme4::lmer(d15n ~ treatment + (1|block), filter(isotopes_e_w_model, sampling == 3))
d15n_s3_null <- lme4::lmer(d15n ~ 1 + (1|block), filter(isotopes_e_w_model, sampling == 3))

summary(d15n_s3_all)
summary(d15n_s3_null)
anova(d15n_s3_all, d15n_s3_null)

d15n_s3_all.emm <- emmeans(d15n_s3_all, pairwise ~ treatment) #, lmer.df="satterthwaite") #method 1
summary(d15n_s3_all.emm)

(s3_d15n <- isotopes_e_w_graph %>% 
    filter(sampling ==3 ) %>% 
    ggplot(aes(x = cage, 
               y = d15n, 
               fill = treatment))+
    geom_boxplot() + 
    geom_jitter(width = 0.1,
                aes(color = cage)) + 
    scale_fill_viridis(name = 'Treatment', discrete = TRUE) +
    facet_grid(treatment ~ block) +
    theme_flat)
```


#### Removing outliers from sampling 3

There was one spider cage that had very high values relative to the other samples in sampling 3 (both raw values were pretty high, same for cage 16 in the sampling 3 control but that one was less extreme relative to the upper IQR). So here are the same analyses with cage 35 removed:

```{r}
isotopes_no_outliers_s3 <- isotopes_e_w_model %>% 
  filter(sampling == 3,
         cage != 35)

d15n_out3_all <- lme4::lmer(d15n ~ treatment + (1|block), isotopes_no_outliers_s3)
d15n_out3_null <- lme4::lmer(d15n ~ 1 + (1|block), isotopes_no_outliers_s3)

anova(d15n_out3_all, d15n_out3_null)

d15n_out3_all.emm <- emmeans(d15n_out3_all, pairwise ~ treatment) #, lmer.df="satterthwaite") #method 1
summary(d15n_out3_all.emm)
```

## total percent N

### boxplot

```{r echo=FALSE, message=FALSE, warning=FALSE}
###### ...Figure for total_percent_n=====

## boxplot of total_percent_n
(total_n_box <- isotopes_e_w_graph %>% 
   ggplot(aes(x = treatment, 
              y = total_percent_n, 
              fill = treatment))+
   geom_boxplot() + 
   geom_jitter(width = 0.1, aes(color = cage)) +
   scale_fill_viridis(name = 'Treatment', discrete = TRUE) +
   labs(title = 'Total percent N of soil beneath litter bags, sampling 2 and 3',
        x = 'Treatment',
        y = 'Total %N',
        caption = 'Possible outliers: samp3, cage 17; samp 3, cage 23') +
   facet_wrap(vars(dir)) +
   theme_flat)

ggplotly(total_n_box)

## checking boxplot outliers function for outliers
out <- boxplot.stats(isotopes_e_w_graph$total_percent_n)$out
out_ind <- which(isotopes_e_w_graph$total_percent_n %in% c(out))
out_ind
isotopes_e_w_graph[out_ind,]

```

the boxplot out function suggests sampling 2, cages 28 + 30 are outliers. Skeptical.


### bar plots
```{r echo=FALSE, message=FALSE, warning=FALSE}
## barplot for mean total_percent_n by treatment and direction
(soil_total_n <- isotopes_e_w_graph %>% 
   group_by(treatment) %>% 
   summarize(mean_total_n = mean(total_percent_n),
             se_total_n = se(total_percent_n)) %>% 
   ungroup() %>% 
   ggplot(aes(x = treatment, y = mean_total_n, fill = treatment))+
   geom_col(position = 'dodge',
            color = 'black')+
   # geom_jitter(data = filter(isotopes_e_w, dir == 'W'),
   geom_jitter(data = isotopes_e_w,
               aes(x = treatment, 
                   y = total_percent_n,
                   group = cage),
               width = 0.2) +
   geom_errorbar(aes( ymin = mean_total_n - se_total_n, 
                      ymax = mean_total_n + se_total_n,
                      width = 0.15)) +
   scale_fill_viridis(name = 'Grashopper fate', discrete = TRUE) +
   labs(title = 'Total %N of soil beneath litter bags, sampling 2 and 3',
        x = 'Treatment',
        y = 'Total %N') +
   # scale_x_discrete(labels=c('Grasshopper', 'Mantid', 'Spider', 'Control')) +
   theme_flat + 
   theme(legend.position = 'none') + 
   facet_wrap(vars(dir))
)
ggplotly(soil_total_n)

# my_ggsave(here::here('results/soil_total_n.png'))
```

### Statistical analysis

#### Sampling 2, from under East litterbags

```{r}
total_n_s2_all <- lme4::lmer(total_percent_n ~ treatment + (1|block), filter(isotopes_e_w_model, sampling == 2))
total_n_s2_null <- lme4::lmer(total_percent_n ~ 1 + (1|block), filter(isotopes_e_w_model, sampling == 2))

anova(total_n_s2_all, total_n_s2_null)

```
Not significant, so no posthoc.


#### Sampling 3, from under West litterbags

```{r}
total_n_s3_all <- lme4::lmer(total_percent_n ~ treatment + (1|block), filter(isotopes_e_w_model, sampling == 3))
total_n_s3_null <- lme4::lmer(total_percent_n ~ 1 + (1|block), filter(isotopes_e_w_model, sampling == 3))

anova(total_n_s3_all, total_n_s3_null)

total_n_s3_all.emm <- emmeans(total_n_s3_all, pairwise ~ treatment) #, lmer.df="satterthwaite") #method 1
summary(total_n_s3_all.emm)

```


#### Removing outliers from sampling 3


```{r}
isotopes_no_outliers_s3 <- isotopes_e_w_model %>% 
  filter(sampling == 3,
         cage != 23)

total_n_out3_all <- lme4::lmer(total_percent_n ~ treatment + (1|block), isotopes_no_outliers_s3)
total_n_out3_null <- lme4::lmer(total_percent_n ~ 1 + (1|block), isotopes_no_outliers_s3)

anova(total_n_out3_all, total_n_out3_null)

total_n_out3_all.emm <- emmeans(total_n_out3_all, pairwise ~ treatment) #, lmer.df="satterthwaite") #method 1
summary(total_n_out3_all.emm)
```



## d13C in the soil

### boxplot
```{r echo=FALSE, message=FALSE, warning=FALSE}
###### ...Figure for d13C=====

## boxplot of d13c
(d13c_box <- isotopes_e_w_graph %>% 
   ggplot(aes(x = treatment, 
              y = d13c, 
              fill = treatment))+
   geom_boxplot() + 
   geom_jitter(width = 0.1,
               aes(color = cage)) + 
   scale_fill_viridis(name = 'Treatment', discrete = TRUE) +
   facet_grid(. ~ dir) +
   theme_flat)

ggplotly(d13c_box)
```

### barplots

```{r echo=FALSE, message=FALSE, warning=FALSE}
## barplot for mean d13C by treatment and direction
(soil_d13c <- isotopes_e_w_graph %>% 
   group_by(treatment) %>% 
   summarize(mean_d13c = mean(d13c),
             se_d13c = se(d13c)) %>% 
   ungroup() %>% 
   ggplot(aes(x = treatment, y = mean_d13c, fill = treatment))+
   geom_col(position = 'dodge',
            color = 'black')+
   # geom_jitter(data = filter(isotopes_e_w, dir == 'W'),
   geom_jitter(data = isotopes_e_w,
               aes(x = treatment, 
                   y = d13c,
                   group = cage),
               width = 0.2) +
   geom_errorbar(aes( ymin = mean_d13c - se_d13c, 
                      ymax = mean_d13c + se_d13c,
                      width = 0.15)) +
   scale_fill_viridis(name = 'Grashopper fate', discrete = TRUE) +
   labs(title = 'd13C of soil beneath litter bags, samplings 2 and 3',
        x = 'Treatment',
        y = 'd13C') +
   # scale_x_discrete(labels=c('Grasshopper', 'Mantid', 'Spider', 'Control')) +
   theme_flat + 
   theme(legend.position = 'none') + 
   facet_wrap(vars(dir))
)

# my_ggsave(here::here('results/soil_d13c.png'))
```


### Statistical analysis

#### Sampling 2 d13C, from under East litterbags

```{r}
d13c_s2_all <- lme4::lmer(d13c ~ treatment + (1|block), filter(isotopes_e_w_model, sampling == 2))
d13c_s2_null <- lme4::lmer(d13c ~ 1 + (1|block), filter(isotopes_e_w_model, sampling == 2))

summary(d13c_s2_all)
summary(d13c_s2_null)
anova(d13c_s2_all, d13c_s2_null)

# not running emmeans because it's not significant
# d13c_s2_all.emm <- emmeans(d13c_s2_all, pairwise ~ treatment) #, lmer.df="satterthwaite") #method 1
# summary(d13c_s2_all.emm)
```
Not significant.


#### Sampling 3, from under West litterbags

```{r}
d13c_s3_all <- lme4::lmer(d13c ~ treatment + (1|block), filter(isotopes_e_w_model, sampling == 3))
d13c_s3_null <- lme4::lmer(d13c ~ 1 + (1|block), filter(isotopes_e_w_model, sampling == 3))

summary(d13c_s3_all)
summary(d13c_s3_null)
anova(d13c_s3_all, d13c_s3_null)

```


#### Removing outliers from sampling 3

There was one spider cage that had very high values relative to the other samples in sampling 3 (both raw values were pretty high, same for cage 16 in the sampling 3 control but that one was less extreme relative to the upper IQR). So here are the same analyses with cage 35 removed:

```{r}
isotopes_no_outliers_s3 <- isotopes_e_w_model %>% 
  filter(sampling == 3,
         cage != 35)

d13c_out3_all <- lme4::lmer(d13c ~ treatment + (1|block), isotopes_no_outliers_s3)
d13c_out3_null <- lme4::lmer(d13c ~ 1 + (1|block), isotopes_no_outliers_s3)

anova(d13c_out3_all, d13c_out3_null)

d13c_out3_all.emm <- emmeans(d13c_out3_all, pairwise ~ treatment) #, lmer.df="satterthwaite") #method 1
summary(d13c_out3_all.emm)
```

## Total percent C

### boxplot

```{r echo=FALSE, message=FALSE, warning=FALSE}
###### ...Figure for total_percent_c=====

## boxplot of total_percent_c
(total_c_box <- isotopes_e_w_graph %>% 
   ggplot(aes(x = treatment, 
              y = total_percent_c, 
              fill = treatment))+
   geom_boxplot() + 
   geom_jitter(width = 0.1, aes(color = cage)) +
   scale_fill_viridis(name = 'Treatment', discrete = TRUE) +
   labs(title = 'Total percent C of soil beneath litter bags, sampling 2 and 3',
        x = 'Treatment',
        y = 'Total %C',
        caption = 'Possible outliers: samp3, cage 17') +
   facet_wrap(vars(dir)) +
   theme_flat)

ggplotly(total_c_box)

## checking boxplot outliers function for outliers
out <- boxplot.stats(isotopes_e_w_graph$total_percent_c)$out
out_ind <- which(isotopes_e_w_graph$total_percent_c %in% c(out))
isotopes_e_w_graph[out_ind,]

```

The boxplot "out" function doesn't suggests any outliers.


### bar plots
```{r echo=FALSE, message=FALSE, warning=FALSE}
## barplot for mean total_percent_c by treatment and direction
(soil_total_c <- isotopes_e_w_graph %>% 
   group_by(treatment) %>% 
   summarize(mean_total_c = mean(total_percent_c),
             se_total_c = se(total_percent_c)) %>% 
   ungroup() %>% 
   ggplot(aes(x = treatment, y = mean_total_c, fill = treatment))+
   geom_col(position = 'dodge',
            color = 'black')+
   # geom_jitter(data = filter(isotopes_e_w, dir == 'W'),
   geom_jitter(data = isotopes_e_w,
               aes(x = treatment, 
                   y = total_percent_c,
                   group = cage),
               width = 0.2) +
   geom_errorbar(aes( ymin = mean_total_c - se_total_c, 
                      ymax = mean_total_c + se_total_c,
                      width = 0.15)) +
   scale_fill_viridis(name = 'Grashopper fate', discrete = TRUE) +
   labs(title = 'total %C of soil beneath litter bags, sampling 2 and 3',
        x = 'Treatment',
        y = 'total %C') +
   # scale_x_discrete(labels=c('Grasshopper', 'Mantid', 'Spider', 'Control')) +
   theme_flat + 
   theme(legend.position = 'none') + 
   facet_wrap(vars(dir))
)
ggplotly(soil_total_c)

# my_ggsave(here::here('results/soil_total_c.png'))
```

### Statistical analysis

#### Sampling 2, from under East litterbags

```{r}
total_c_s2_all <- lme4::lmer(total_percent_c ~ treatment + (1|block), filter(isotopes_e_w_model, sampling == 2))
total_c_s2_null <- lme4::lmer(total_percent_c ~ 1 + (1|block), filter(isotopes_e_w_model, sampling == 2))

anova(total_c_s2_all, total_c_s2_null)

```
Not significant, so no posthoc.


#### Sampling 3, from under West litterbags

```{r}
total_c_s3_all <- lme4::lmer(total_percent_c ~ treatment + (1|block), filter(isotopes_e_w_model, sampling == 3))
total_c_s3_null <- lme4::lmer(total_percent_c ~ 1 + (1|block), filter(isotopes_e_w_model, sampling == 3))

anova(total_c_s3_all, total_c_s3_null)
```
Not significant, so no posthoc.


#### Removing outliers from sampling 3


```{r}
isotopes_no_outliers_s3 <- isotopes_e_w_model %>% 
  filter(sampling == 3,
         cage != 17)

total_c_out3_all <- lme4::lmer(total_percent_c ~ treatment + (1|block), isotopes_no_outliers_s3)
total_c_out3_null <- lme4::lmer(total_percent_c ~ 1 + (1|block), isotopes_no_outliers_s3)

anova(total_c_out3_all, total_c_out3_null)

total_c_out3_all.emm <- emmeans(total_c_out3_all, pairwise ~ treatment) #, lmer.df="satterthwaite") #method 1
summary(total_c_out3_all.emm)
```

Weakly significant without cage 17, though posthoc is not significant.








[[---------for analyses until 2021.09.21----------]]


```{r}
## data analysis: trying anovas for d13c ----

## checking the overall model, with all the data 
d15n_anova_all <- aov(d15n ~ treatment + block + dir, isotopes_e_w_model)
summary(d15n_anova_all)

d15n_anova_simple <- aov(d15n ~ treatment, isotopes_e_w_model)
summary(d15n_anova_simple)

anova(d15n_anova_simple, d15n_anova_all)
```

The effect is weaker/non-existent if we separate it by direction (lmer has singularity issues with random effects). E soil samples have a weak effect, while W has none.

**The posthoc analysis needs fixing--- but it looks like the main difference is between mantid and spider treatments**

```{r}
##  simple ANOVA and one lmer model
d15n_anova_block_e <- aov(d15n ~ treatment + block, filter(isotopes_e_w_model, dir == 'E'))
summary(d15n_anova_block_e)

d15n_anova_block_w <- aov(d15n ~ treatment + block, filter(isotopes_e_w_model, dir == 'W'))
summary(d15n_anova_block_w)

d15n_lm <- lmer(d15n ~ treatment + (1|block), filter(isotopes_e_w_model, dir == 'W'))
summary(d15n_lm)

## posthoc analysis for E soil samples (only model that was borderline significant)
emmeans(d15n_anova_block_e, pairwise ~ treatment) #, lmer.df="satterthwaite") #method 1

# TukeyHSD(d15n_anova_block_e) ## doesn't work

isotopes_e <- filter(isotopes_e_w_model, dir == 'E')
pairwise.t.test(isotopes_e$d15n, isotopes_e$treatment, p.adj = 'bonf') ## adjustments include bonf, holm

```


```{r}
## Data analysis: d13C by treatment ----
## remember to use isotope_e_w_model for comparisons against the control

# isotopes_e_w <- mutate(isotopes_e_w,
#                        d15n_2 = scale(d15n, center = TRUE, scale = TRUE))

lm_soil_d15n <- lme4::lmer(d15n ~ treatment + (1|block), isotopes_e_w_model)
lm_soil_d15n_null <- lme4::lmer(d15n ~ (1|block), isotopes_e_w_model)

## Nevo's method of determining whether a variable matters and the post hoc
anova(lm_soil_d15n, lm_soil_d15n_null)
emmeans(lm_soil_d15n, pairwise ~ treatment)

# # old code
# summary(lm_soil_d15n)
# emmeans(lm_soil_d15n, pairwise ~ treatment) #, lmer.df="satterthwaite") #method 1

par(mfrow = c(2,2))
plot(lm_soil_d15n)

```

If pooling the data between sampling 2 and sampling 3, treatment is significant. Primarily, control and spider differ, and mantid and spider differ. 


If looking ONLY at the sampling 3 data...

```{r}
## Data analysis: d13C by treatment ----
isotopes_w <- filter(isotopes_e_w_model, dir == 'W')

lm_soil_d15n <- lme4::lmer(d15n ~ treatment + (1|block), isotopes_w)
lm_soil_d15n_null <- lme4::lmer(d15n ~ (1|block), isotopes_w)

## Nevo's method of determining whether a variable matters and the post hoc
anova(lm_soil_d15n, lm_soil_d15n_null)
emmeans(lm_soil_d15n, pairwise ~ treatment)

# # old code
# summary(lm_soil_d15n)
# emmeans(lm_soil_d15n, pairwise ~ treatment) #, lmer.df="satterthwaite") #method 1

par(mfrow = c(2,2))
plot(lm_soil_d15n)

```

Thoughts from organizing code 2021.09.18:
The results I found in the spring were based on pooled data from sampling 2 and 3. I'm not totally sure how legit that is, and when the data are separated, there is barely any significance of an effect. 

As for the more complete statistical analyses, we see basically the same thing. Only significant when pooled, and there are issues with singularity when block is a random effect.


After talking to Dror 2021.09.19:
```{r}
## Data analysis: d13C by treatment ----
isotopes_w <- filter(isotopes_e_w_model, dir == 'W')
isotopes_e <- filter(isotopes_e_w_model, dir == 'E')

lm_soil_d15n_s2 <- lmer(d15n ~ treatment + (1|block), isotopes_e)
summary(lm_soil_d15n_s2)
lm_soil_d15n_s3 <- lmer(d15n ~ treatment + (1|block), isotopes_w)
summary(lm_soil_d15n_s3)

par(mfrow = c(2,2))
plot(lm_soil_d15n_s3)

```

