---
title: 'Exp 3 Results: Plant Isotopes at Experiment End'
author: "Y Tang"
date: "19 August 2021"
output: 
  html_document: 
    toc: true
    toc_depth: 4
    toc_float: true
    theme: united
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(knitr)
library(lme4)
library(lmerTest)
library(viridis)
library(emmeans)

source(here::here('src/yaya_fxns.R'))
# imports isotopes_plants_long
source(here::here('src/g9.1_isotope_plant_end_import.R'))
# theme objects are in yaya_fxns.R


isotopes_plants_long <- isotopes_plants_long %>%
  mutate(treatment = factor(treatment, levels = c('ghop', 'mantid', 'spider','control')),
         fxl_group = factor(fxl_group, levels = c('cer', 'leg', 'C.a.', 'ot')))
```

## Aboveground plant elemental + isotopic data

We analyzed the 13C and 15N content of homogenized plant matter at the end of the experiment, categorized by functional group. First, lets explore the data distribution.

These are distributions of %C, d13C, %N, and d15N.

(**these graphs can be compressed into a small, four panel table**)

```{r}
# for generating histograms of initial data analysis
explore_histogram <- function(var_name) {
  
  isotopes_plants_long %>% 
    filter(analysis == var_name) %>% 
    ggplot(aes(a_value))+
    geom_histogram(bins = 10) + 
    labs(title = paste('Histogram of', var_name, 'values'),
         x = var_name) +
    facet_grid(fxl_group ~ treatment) + 
    theme_light()
}

par(mar = c(4, 4, .1, .1))

explore_histogram('total_percent_c')
explore_histogram('d13c')
explore_histogram('total_percent_n')
explore_histogram('d15n')

```


Here is a plot of paired variables:

```{r}

isotopes_plants_long %>% 
  pivot_wider(names_from = analysis, values_from = a_value) %>%
  select(c('total_percent_n', 'd15n','atom_percent_15n', 'total_percent_c', 'd13c','atom_percent_13c')) %>% 
  pairs()

```

### Elemental data plots

Now lets take a look at the results of elemental analysis between treatments.

```{r}
# isotopes_plants_summ <-
isotopes_plants_long %>%
  # group_by(treatment, analysis, fxl_group) %>%
  group_by(treatment, analysis) %>%
  summarize(iso_mean = mean(a_value),
            iso_se = se(a_value),
            iso_var = var(a_value),
            iso_n = n()) %>% 
  ggplot(aes(fill = treatment, 
             x = analysis, 
             y = iso_mean))+
  geom_col(position = 'dodge') + 
  # geom_bar(position = 'dodge', stat = 'identity')+
  # geom_point(data = isotopes_plants_long,
  #            aes(x = fxl_group,
  #                y = a_value,
  #                color = treatment),
  #            position = position_jitterdodge()) +
  geom_errorbar(aes(ymin=iso_mean-iso_se, ymax= iso_mean+iso_se), width=.2,
                position=position_dodge(.9)) +
  scale_fill_viridis(name = 'Grasshopper fate', discrete = TRUE) +
  labs(title = 'Comparing all elemental analyses by treatment',
       x = 'Analysis',
       y = 'Mean') + 
  # facet_wrap(vars(analysis), scales = 'free') +
  theme_angled

## graphing by treatment + fxl_group, faceted by the analysis variable
isotopes_plants_long %>%
  # pivot_wider(names_from = 'analysis', values_from = 'a_value') %>% 
  # group_by(treatment, fxl_group) %>%
  # View()
  # summarize(iso_mean = mean(a_value),
  #           iso_se = se(a_value),
  #           iso_var = var(a_value),
  #           iso_n = n()) %>% 
  ggplot(aes(fill = fxl_group,
             x = treatment,
             y = a_value))+
  geom_boxplot() + 
  # geom_bar(position = 'dodge', stat = 'identity')+
  # geom_errorbar(aes(ymin=iso_mean-iso_se, ymax= iso_mean+iso_se), width=.2,
  # position=position_dodge(.9)) +
  scale_fill_viridis(discrete = TRUE, option = 'magma') +
  labs(title = 'Comparing all elemental analyses by treatment') + 
  facet_grid(treatment*fxl_group ~ analysis, scales = 'free') +
  theme_angled


```

This is for d15N:

```{r}
plot_analysis <- function(analysis_name) {
  
  isotopes_plants_long %>% 
    filter(analysis == analysis_name) %>%
    group_by(treatment) %>% 
    mutate(treatment = factor(treatment, levels = c('ghop', 'mantid', 'spider','control'))) %>% 
    summarize(mean_analysis = mean(a_value),
              se_analysis = se(a_value)) %>% 
    ggplot(aes(x = treatment, y = mean_analysis, fill = treatment))+
    geom_col(position = 'dodge',
             color = 'black')+
    geom_errorbar(aes( ymin = mean_analysis - se_analysis, 
                       ymax = mean_analysis + se_analysis,
                       width = 0.15)) +
    scale_fill_viridis(name = 'Treatment', discrete = TRUE) +
    labs(title = paste(analysis_name, 'in plant tissue at end of experiment'),
         x = 'Treatment',
         y = analysis_name) +
    # scale_x_discrete(labels=c('Grasshopper', 'Mantid', 'Spider', 'Control')) +
    theme_light()
}

anova_analysis <- function(analysis_name) {
  sub_anova <- aov(a_value ~ treatment + block, data = filter(isotopes_plants_long, analysis == analysis_name))
  print(summary(sub_anova))
}

lmer_analysis <- function(analysis_name) {
  sub_lmer <- lme4::lmer(a_value ~ treatment + block, data = filter(isotopes_plants_long, analysis == analysis_name))
  print(summary(sub_lmer))
}


plot_analysis('d15n')
# sub_anova <- aov(a_value ~ treatment, data = filter(isotopes_plants_long, analysis == 'd15n'))

anova_analysis('d15n')


plot_analysis('total_percent_c')
anova_analysis('total_percent_c')

plot_analysis('d13c')
anova_analysis('d13c')

plot_analysis('total_percent_n')
anova_analysis('total_percent_n')



```

These are the statistical models for each analysis.

```{r}

isotopes_plants_long$treatment <- fct_relevel(isotopes_plants_long$treatment, c('control', 'ghop', 'mantid', 'spider')) 

all_vars <- unique(isotopes_plants_long$analysis)
var_name <- all_vars[2] # 1: %n, 2: d15n, 4: total %c, 5: d13c

lmer_plant_analysis <- 
  lm(a_value ~ treatment + (1|block),
     filter(isotopes_plants_long, analysis == var_name))

summary(lmer_plant_analysis)
# emmeans(lmer_plant_analysis, pairwise ~ treatment) #, lmer.df="satterthwaite") #method 1

## check the residuals 
library(ggpubr)
ggqqplot(resid(lmer_plant_analysis))

# this is a simple ANOVA, without fxl groups or blocks
anov_plant_analysis <- 
  aov(a_value ~ treatment, #+ block,
      filter(isotopes_plants_long, analysis == var_name))
summary(anov_plant_analysis)

```

After running analyses on each of the four analysis variables, none had a significant overall effect of treatment.



Lets try out ART anova....

```{r}
library(ARTool)

model_var <- 'total_percent_n'
artanova_model <- art(a_value ~ treatment + fxl_group + treatment*fxl_group + (1|block),
                      filter(isotopes_plants_long, analysis == model_var))
anova(artanova_model)
```

It agrees.

Now lets try it with the functional group as a predictor variable, because it may be that there is no OVERALL difference between treatments, but specific subgroups within the plant community benefit more from one predator treatment over another.

This is the variable that is being compared:

```{r}
## all_vars is made of 
# [1] total % n
# [2] d15n
# [3] atom % 15n
# [4] total % c
# [5] d13c
# [6] atom % 13c
(var_name <- all_vars[1])

## linear mixed model, with blocking as random effect
lmer_plant_analysis <- 
  lm(a_value ~ treatment + fxl_group + treatment*fxl_group,
     filter(isotopes_plants_long, analysis == var_name))

summary(lmer_plant_analysis)
emmeans(lmer_plant_analysis, pairwise ~ fxl_group) #, lmer.df="satterthwaite") #method 1

## check the residuals 
library(ggpubr)
# ggqqplot(resid(lmer_plant_analysis))

```

The results here are for the model `a_value ~ treatment + fxl_group + treatment*fxl_group`, without blocking (as of 2021.08.19 I am running it without blocking, due to singularity issues).

%N: 
We see a significant effect of N in legumes, which differs from C.arg (the "control" in this model), cereals, and others. (slope 1.0175, p << 0.001). This makes sense. No effect of treatment.

d15N 
Same results as above. 

%C
Also higher plant %C in legumes (p < 0.05).

d13C 
Cereals + legumes differ significantly from C.arg and others.



## Isotope scatterplot hair graph

Lets look at the isotopic separation for funsies.

```{r}
iso_summ <- isotopes_plants_long %>%
  pivot_wider(names_from = analysis, values_from = a_value) %>%
  group_by(treatment, fxl_group) %>%
  summarize(mean_d15n = mean(d15n),
            se_d15n = se(d15n),
            mean_d13c = mean(d13c),
            se_d13c = se(d13c),
            mean_percent_n = mean(total_percent_n),
            se_percent_n = se(total_percent_n),
            mean_percent_c = mean(total_percent_c),
            se_percent_c = se(total_percent_c),
            n_size = n())

iso_summ %>% 
  ggplot(aes(x = mean_d13c, 
             y = mean_d15n, 
             color = treatment))+
  geom_point(aes(shape = factor(fxl_group)),
             size = 2) +
  geom_errorbar(width = 0.04, aes(ymin = mean_d15n - se_d15n, ymax = mean_d15n + se_d15n))+
  geom_errorbarh(height = 0.5, aes(xmin = mean_d13c - se_d13c, xmax = mean_d13c + se_d13c))+
  labs(title = "Trophic position plot (?)")


```

