---
title: "Exp 3: Litterbag and ash-free analysis"
author: "Y Tang"
date: "18 September 2021"
output: 
  html_document: 
    toc: true
    toc_depth: 4
    toc_float: true
    theme: united
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#=============
library(tidyverse)
library(plotly)
library(janitor)
library(ggpubr)
library(viridis)
library(here)
library(emmeans)
source(here::here('src/yaya_fxns.r'))

theme_ash <- function(){  # Creating a function
  theme_classic() +  # Using pre-defined theme as base
    theme(axis.text.x = element_text(size = 12, face = "bold"),  # Customizing axes text      
          axis.text.y = element_text(size = 12, face = "bold"),
          axis.title = element_text(size = 14, face = "bold"),  # Customizing axis title
          panel.grid = element_blank(),  # Taking off the default grid
          plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), units = , "cm"),
          legend.text = element_text(size = 12, face = "italic"),  # Customizing legend text
          legend.title = element_text(size = 12, face = "bold"),  # Customizing legend title
          legend.position = "right",  # Customizing legend position
          plot.caption = element_text(size = 12))  # Customizing plot caption
}                                  

graph_levels <- c('ghop', 'mantid', 'spider', 'control')
model_levels <- c('control', 'ghop', 'mantid', 'spider')
```


# Measuring decompositon

First I wanted to see whether there was evidence of decomposition at the end of the experiment. This requires the dry mass data from W and S litterbags, which were initially weighed by Evya in April 2020, cleaned and re-weighed by Yaya in May 2020. Litterbag data used here is from September 2021, when litter was removed from litterbags and weighed directly.

Subsamples of litter from each litterbag were ashed at 550C for five hours. All litter data were analyzed on an ash-free basis to remove the effect of mineral soil that may have entered the litterbags. 

```{r message=FALSE, warning=FALSE, , message=FALSE}
## import litterbag litter only data, litterbag ashed data, and the cage treatment data

litter_only_raw <- read_csv(here::here('data/litterbag_litter_only_DATA_2021.09.csv'))%>% 
  clean_names() %>% 
  mutate(cage = as.numeric(substr(sample_id, 0, 2)),
         dir = factor(substr(sample_id, 3, 3)),
         sampling = ifelse(dir == 'E', 2, 3))

# ashed data including the litter from the start of the experiment
ashed_raw <- read_csv(here::here('data/litterbag_ash-free_analysis_samp2_samp3_2021.09.csv')) %>% 
  clean_names() %>% 
  mutate(cage = as.numeric(substr(sample_id, 0, 2)),
         dir = factor(substr(sample_id, 3, 3)))

cage_treatments <- read_csv(here::here('results/g1_updated_cages_wk9.csv')) %>% 
  select(-predatorID, -position)

# import experimental start litter mass
litter_start_raw <- read_csv(here::here('data/litterbag_DATA_2019.07.11.csv')) %>% 
  clean_names() %>% 
  rename(dir = posit,
         litter_start_mass = litter_mass) %>% 
  filter(cage != 3,
         cage != 26) #cages that were destroyed early on

#join litter data with cage info
litter_only_data <- left_join(litter_only_raw, cage_treatments) %>%
  mutate(treatment = fct_relevel(treatment, c('control', 'ghop', 'mantid', 'spider')))

ash_free_calc <- left_join(ashed_raw, litter_only_data) %>% 
  mutate(treatment = fct_relevel(treatment, c('control', 'ghop', 'mantid', 'spider')),
         sample_to_ash = pre_oven_dry_mass - crucible_weight,
         litter_ash = post_oven_ash_mass - crucible_weight,
         ash_proportion = litter_ash / sample_to_ash,
         litter_corrected = total_litterbag_litter * (1 - ash_proportion)) %>% 
  select(-crucible_weight, -pre_oven_dry_mass, -post_oven_ash_mass)

ash_free_data <- ash_free_calc %>% 
  filter(!is.na(cage)) #,
# cage != 40,
# cage != 2) ## cage 2 had outlier values of corrected litter (too high, not much decomp overall)

# ashed values for litter at start
ashed_litter_start <- ash_free_calc %>% 
  filter(dir == 'i',
         sample_id != 'Mlit-Pow') %>%  ## powder ash proportion was higher than litter
  select(sample_id, sample_to_ash, litter_ash, ash_proportion)

litter_start_data <-  litter_start_raw %>% 
  mutate(ash_proportion = mean(ashed_litter_start$ash_proportion),
         litter_start_corr_mass = litter_start_mass * (1-ash_proportion),
         sampling = 0) %>% 
  select(-closed_bag_litter) %>% 
  left_join(unique(select(litter_only_data, c('cage', 'dir', 'block', 'treatment', 'replicate')))) %>% 
  drop_na()

```


## Destructively sampled litterbag litter properties

Sampling 2 bags are "East", and sampling 3 from the end of the experiment are the "South" and "West" bags.

These data are for litterbags destructively sampled from cages at various points in the experiment. 

This is a histogram of the litter mass prior to ashing, for all three litterbag directions (coloring to give a sense of the distribution within each litterbag direction).


### Histogram of uncorr. litter, all samples together (2 + 3, E-S-W)

Uncolored histogram of litterbag litter mass.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# hist(litter_only_data$total_litterbag_litter)
litter_only_data %>% 
  ggplot(aes(x = total_litterbag_litter))+
  geom_histogram(bins = 10)+
  theme_ash()
```


### Histogram of all uncorr. litter, by direction


```{r message=FALSE, warning=FALSE, message=FALSE}
litter_only_data %>% 
  ggplot(aes(x = total_litterbag_litter)) + 
  geom_histogram(bins = 10) +
  facet_grid(dir  ~ .)+
  theme_ash()
```


### Litterbag litter by cage and treatment, across entire experiment

This is the uncorrected mass of the litter from the start of the experiment through samplings 2 and 3 (only explicitly weighed litterbags) for each cage (replicate x treatment):

```{r echo=FALSE, message=FALSE, warning=FALSE}
litter_start_data %>% ## the raw df has the initial mass
  rename(total_litterbag_litter = litter_start_mass) %>% 
  mutate(sampling = 0) %>% 
  bind_rows(litter_only_data) %>% 
  ggplot(aes(x = cage,
             y = total_litterbag_litter,
             color = dir))+
  geom_point(aes(shape = treatment)) +
  scale_color_viridis(discrete = TRUE) + 
  expand_limits(y = 0) + 
  facet_wrap(vars(sampling)) +
  theme_ash()

# facet_grid(replicate ~ treatment) +
#, direction = -1)

litter_start_data %>% ## the raw df has the initial mass
  rename(total_litterbag_litter = litter_start_mass) %>% 
  mutate(sampling = 0,
         treatment = factor(treatment, graph_levels)) %>% 
  bind_rows(litter_only_data) %>% 
  ggplot(aes(x = sampling,
             y = total_litterbag_litter))+
  geom_col(aes(fill = treatment)) +
  scale_fill_viridis(discrete = TRUE) + #, direction = -1) + 
  expand_limits(y = 0) + 
  facet_grid(replicate ~ treatment) + ## don't use block, since there are two controls in block 10 and it doubles the values
  theme_ash() 
```


## Ash-free data properties

### Histogram of ash-free proportion, all samples

All samples together, and then by sampling direction.

```{r message=FALSE, warning=FALSE, message=FALSE}
ash_free_data %>% 
  ggplot(aes(x = ash_proportion)) + 
  geom_histogram(bins = 10)+
  theme_ash()

## plotting some ash free data properties
ash_free_data %>% 
  ggplot(aes(x = ash_proportion)) +
  geom_histogram(bins = 10) + 
  facet_grid(dir ~ .) +
  theme_ash()
```

### Histogram of ash-free proportion, samp 3 only (W + S)

```{r message=FALSE, warning=FALSE, , message=FALSE}
# spread of corrected litterbag masses
ash_free_data %>% 
  filter(dir != 'E') %>%
  ggplot(aes(x = litter_corrected)) + 
  # fill = treatment)) + 
  geom_histogram(bins = 10) + 
  scale_fill_viridis(discrete = TRUE) +
  theme_ash()
```


Lets take a look a the _corrected_ masses of the litterbag litter.

Corrected litterbag litter is calculated as the difference between the total litterbag mass minus the ash proportion calculated from the ashed subsample.


```{r message=FALSE, warning=FALSE, message=FALSE}
## Creating a boxplot showing the two samplings of corrected litterbag ash-free mass

ash_free_data %>%
  ggplot(aes(x = treatment,
             y = ash_proportion)) +
  geom_boxplot() + 
  geom_jitter(width = 0.2,
              aes(color = dir)) +
  scale_color_viridis(discrete = TRUE) + 
  # scale_color_viridis(name = 'Treatment', labels = c('Control', 'Carcass', 'Mantid', 'Spider'), discrete = TRUE) +
  labs(title = "Ash proportion by sampling",
       x = "\nTreatment",
       y = "Ash proportion\n",
       caption = 'Facets are for each of the sampling events (2 and 3).') +
  expand_limits(y=0) +
  facet_wrap(vars(sampling)) +
  theme_ash()
```

This is the plot of the correct litter masses, from sampling 0 to 2 and 3. 

```{r message=FALSE, warning=FALSE, message=FALSE}
# Creating a boxplot with ggplot for looking at corrected litter amounts
litter_boxplot_dir <- ash_free_data %>% 
  bind_rows(rename(litter_start_data, litter_corrected = litter_start_corr_mass)) %>% 
  ggplot(aes(x = treatment,
             y = litter_corrected,
             fill = dir,
             group = cage)) +
  geom_boxplot() +
  geom_jitter() + 
  scale_fill_viridis(discrete = TRUE) +
  # scale_fill_manual(values = c("#97F7C5", "#4ED973", "#08873D")) +
  labs(x = "\nTreatment\n", 
       y = "\nCorrected litter mass\n",
       caption = "\n Fig.2 Mass of litter in litterbags at the end of the
                        experiment.") +
  # caption = "\nFig.2") +  # Caption for figure in panel
  expand_limits(y = 0) +
  facet_wrap(vars(sampling)) +
  theme_ash()

ggplotly(litter_boxplot_dir)
```

For reference, this is the ash proportion of the litter at the start (before installation into cages):

```{r message=FALSE, warning=FALSE, message=FALSE}
ashed_litter_start %>% 
  summarize(mean_ash = mean(ash_proportion)) %>% 
  knitr::kable()

```

This table is the average ash proportion for each direction x treatment.

```{r message=FALSE, warning=FALSE, message=FALSE}
ash_free_data %>% 
  group_by(dir, treatment) %>% 
  summarize(mean_ash_prop = mean(ash_proportion),
            se_ash_prop = se(ash_proportion),
            sd_ash_prop = sd(ash_proportion)) %>% 
  knitr::kable()

```

The previous plot that includes sampling 2 and 3 shows some outlier behavior from cage 2 (low decomposition?). They are removed for the next set of graphs. 

This graph shows the spread of corrected litter, from the beginning to samplings 2 and 3.

```{r echo=FALSE}

ash_free_clean <- ash_free_data %>%
  filter(cage != 2) 

plotly_ash_free_clean <- ash_free_clean %>% 
  ggplot(aes(x = treatment,
             y = litter_corrected)) +
  geom_boxplot() + 
  geom_jitter(width = 0.2,
              aes(color = dir)) + 
  scale_color_viridis(discrete = TRUE) +
  # scale_color_viridis(name = 'Treatment', labels = c('Control', 'Carcass', 'Mantid', 'Spider'), discrete = TRUE) +
  labs(title = "Corrected litter masses",
       x = "\nTreatment",
       y = "Corrected litter\n\n",
       caption = 'Boxplots are of all datapoints, while jittered point show the litterbag source') +
  expand_limits(y=0) +
  facet_wrap(vars(sampling)) +
  theme_ash()

ggplotly(plotly_ash_free_clean)
```

## Percent mass loss

After adjusting the dry litter mass with ash-free corrections, here is the percent mass loss for samplings 2 and 3: (with cage 2 removed, because it was an outlier for low decomposition in all bags)

```{r message=FALSE, warning=FALSE, message=FALSE}
percent_decomp <- ash_free_data %>% 
  left_join(select(litter_start_data, -c('sampling', 'block', 'treatment', 'ash_proportion', 'litter_start_mass')), 
            by = c('cage', 'dir')) %>% 
  filter(cage != 2) %>% 
  mutate(percent_mass_loss = (litter_start_corr_mass - litter_corrected)/ litter_start_corr_mass)

percent_decomp %>% 
  group_by(sampling, treatment) %>% 
  summarize(mean_percent_mass_loss = mean(percent_mass_loss),
            se_percent_mass_loss = se(percent_mass_loss)) %>%
  knitr::kable()

p <- percent_decomp %>% 
  ggplot(aes(x = treatment,
             y = percent_mass_loss,
             group = cage)) +
  geom_boxplot() + 
  geom_jitter(width = 0.2,
              aes(color = dir)) + 
  # scale_color_viridis(discrete = TRUE) +
  scale_color_viridis(name = 'Litterbag\ndirection', discrete = TRUE) +
  labs(title = "Corrected litter masses",
       x = "\nTreatment",
       y = "Percent mass loss\n\n") +
  expand_limits(y=0) +
  scale_y_continuous(labels = scales::percent) + 
  facet_wrap(vars(sampling)) +
  theme_ash()

ggplotly(p)
```

And again, with sampling 3 separated out.

```{r}
plot_percent_decomp <- 
  percent_decomp %>% 
  filter(sampling != 2) %>%
  ggplot(aes(x = treatment,
             y = percent_mass_loss,
             group = cage)) +
  geom_boxplot() + 
  geom_jitter(width = 0.2,
              aes(color = dir)) + 
  # scale_color_viridis(discrete = TRUE) +
  scale_color_viridis(name = 'Litterbag\ndirection', discrete = TRUE, option = 'turbo') +
  labs(title = "Corrected litter masses",
       x = "\nTreatment",
       y = "Percent mass loss\n\n") +
  expand_limits(y=0) +
  scale_y_continuous(labels = scales::percent) + 
  facet_wrap(vars(dir)) +
  theme_ash()

ggplotly(plot_percent_decomp)
```

There are no signs of a treatment difference.



## Statistical analysis for percent mass loss 


### ANOVA with blocks as a random effect


The mixed model used here (one-way ANOVA with blocking as a random effect) is:
percent_mass_loss ~ treatment | (1|block)


#### Sampling 2: East

```{r echo=FALSE, message=TRUE, warning=FALSE}
decomp_s2_anova_full <- lme4::lmer(percent_mass_loss ~ treatment + (1|block), 
                                   filter(percent_decomp, sampling == 2), 
                                   REML = FALSE)
# summary(decomp_s2_anova_full)
decomp_s2_anova_null <- lme4::lmer(percent_mass_loss ~ 1 + (1|block), 
                                   filter(percent_decomp, sampling == 2),
                                   REML = FALSE)
# summary(decomp_s2_anova_null)

anova(decomp_s2_anova_full, decomp_s2_anova_null) %>% 
  knitr::kable()

```

REML = FALSE (so ML, maximum likelihood was used, which is what's needed when the fixed effect structure is being tested)

So here's the report from the selected model:
```{r echo=FALSE, message=TRUE, warning=TRUE}
decomp_s2_anova_null_REML <- lme4::lmer(percent_mass_loss ~ 1 + (1|block), 
                                   filter(percent_decomp, sampling == 2),
                                   REML = TRUE)
summary(decomp_s2_anova_null_REML)

## just to check....
# summary(decomp_s2_anova_full)
```



#### Sampling 3: West

```{r echo=FALSE, message=TRUE, warning=FALSE}
## full model with treatment and blocking as random effect
decomp_s3_anova_full <- lme4::lmer(percent_mass_loss ~ treatment + (1|block), 
                                   filter(percent_decomp, sampling == 3))
# summary(decomp_s3_anova_full)
decomp_s3_anova_null <- lme4::lmer(percent_mass_loss ~ 1 + (1|block), 
                                   filter(percent_decomp, sampling == 3))
# summary(decomp_s3_anova_null)

anova(decomp_s3_anova_full, decomp_s3_anova_null) %>% 
  knitr::kable()


# check the residuals of the full model for sampling 2
plot(decomp_s2_anova_full)

# check the residuals of the full model for sampling 3
plot(decomp_s3_anova_full)

```

___CONCLUSIONS____ 
There was no effect of treatment on decomposition rates of litterbags in these cages. 

